### 2025-05-22 paper/main.tex
Universal commits presents...

In all cases, the right‐hand plots (showing $\Delta_k \cdot k^2$) exhibit a roughly $\mathcal O(1 / k^2)$ decay, confirming that the
incremental effect of adding one more sample diminishes proportionally to $1 / k$ as the dataset grows.

@@

Moreover, in addition to determining the dataset size $K$, the procedure also yields a fully trained model as a by‐product of the
iterative updates.

>>>>

In all cases, the right‐hand plots (showing $\Delta_k \cdot k^2$) exhibit a roughly $\mathcal O(1 / k^2)$ decay, confirming that the
incremental effect of adding one more sample diminishes proportionally to $1 / k^2$ as the dataset grows.

@@

Moreover, in addition to determining the dataset size $K$, the procedure also yields a fully trained model as a by‐product of the
iterative updates. Optionally it is possible to reduce $D$ if calculations are too complex, but it will affect the correlation with loss
function.

### 2025-05-22 slides/inn_prac/iteration_2/innprac_2.tex
Universal commits presents...

\documentclass{beamer}
\beamertemplatenavigationsymbolsempty
\usecolortheme{beaver}
\setbeamertemplate{blocks}[rounded=true, shadow=true]
\setbeamertemplate{footline}[page number]
%
\input{new_commands}
%
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amssymb,amsfonts,amsmath,mathtext}
\usepackage{subfig}
\usepackage[all]{xy}
\usepackage{array}
\usepackage{multicol}
\usepackage{hyperref}
\usepackage{hhline}
\graphicspath{{fig/}{../fig/}}

%----------------------------------------------------------------------------------
\title[Loss Landscape Convergence]{Neural Networks Loss Landscape Convergence in Hessian Low-Dimensional Space}
\author[Tem Nikitin \textit{et al.}]{Tem Nikitin \and Nikita Kiselev \and Andrey Grabovoy \and Vladislav Meshkov}
\institute{Moscow Institute of Physics and Technology}
\date{2025}
%----------------------------------------------------------------------------------
\begin{document}

@@

\begin{frame}{Goals}
    \begin{itemize}
        \item Study how neural network loss landscape changes with dataset size
        \item Define and measure $\Delta_k = \mathbb{E} \bigl( \cL_{k+1}(\bw) - \cL_k(\bw) \bigr)^2$
        \item Develop Hessian-based low-dimensional projection method
        \item Derive theoretical bound for $\Delta_k$ via top-$d$ eigenvalues
        \item Validate threshold $k^*$ beyond which further data yield negligible change
    \end{itemize}
\end{frame}

\begin{frame}{One-slide}
    \begin{columns}[c]
        \hspace{-1.5cm}
        \column{0.6\textwidth}
        \includegraphics[width=1.2\textwidth]{img/loss_eigen_-1_individ.pdf}
        \column{0.5\textwidth}
        \begin{itemize}
            \item Hessian projection onto top-$d$ eigenvectors
            \item Monte Carlo estimate of $\Delta_k^{emp} = \mathbb{E} \Bigl( \cL_{k+1} - \cL_{k} \Bigr)^2.$
            \item Analytical bound: $\Delta_k^{th} \ge \Delta_k^{emp}$
            \item Detect dataset threshold $k^*$
        \end{itemize}
    \end{columns}
    \bigskip
    \textbf{Main message:} Loss landscape stabilizes after sufficient sample size.
\end{frame}

@@

\begin{frame}{Problem Statement: Quality Criteria}
    \begin{itemize}
        \item Convergence rate: $\Delta_k = O(1/k)$
        \item Theoretical bound via top-$d$ eigenvalues upper-bounds empirical $\Delta_k$
        \item Plateau in eigenvalue differences $\lambda_i^{k+1}-\lambda_i^k$ indicates threshold
    \end{itemize}
\end{frame}

\begin{frame}{Problem Solution: Theoretical Analysis}
    \begin{columns}[c]
        \column{0.4\textwidth}
        \begin{itemize}
            \item Project parameters: $\bw = \bw^* + \bP \btheta$, $\bP = [\be_1, \dots, \be_d]$
            \item Taylor approx: $\cL_k (\bw^* + \bP \btheta) \approx \cL_k (\bw^*) + \tfrac12 \btheta^{\T} \bLambda_k \btheta$
                  \bigskip
                  \bigskip
            \item Bound:
        \end{itemize}
        \column{0.6\textwidth}
        \includegraphics[width=1.0\textwidth]{img/eigenvalues.pdf}
        \vspace{0.5cm}\hspace{1.8cm}\raggedright{\tiny Figure 1: Eigenvalue decay (p.4)}
    \end{columns}
    $$
        \Delta_k \approx
        \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{i=1}^d (\lambda_{k+1}^i - \lambda_k^i)^2
        + \Bigl( \sum_{i=1}^d (\lambda_{k+1}^i - \lambda_k^i) \Bigr)^2 \Biggr).
    $$
\end{frame}

\begin{frame}{Goals of Computational Experiment}
    \begin{itemize}
        \item Datasets: MNIST, Fashion-MNIST (60k train, 10k test)
        \item MLP: 2 hidden layers, $\sim 10^5$ parameters
        \item Subspace dimension $d=10$, Monte Carlo samples $K=64$, $\sigma=1$
        \item Compare empirical $\Delta_k$ vs theoretical bound across $k$
    \end{itemize}

    \begin{figure}
        \centering
        \includegraphics[width=1.05\textwidth,trim=0 17 0 1.5cm,clip]{img/delta_eigen_1_10_64.pdf}
        \caption*{\tiny Figure 8: Monte Carlo $\Delta_k$ vs $k$ and $\Delta_k\cdot k^2$ (p.11)}
    \end{figure}
\end{frame}

@@

    \begin{figure}
        \centering
        \includegraphics[width=0.7\textwidth]{img/delta_border_1_10_64.pdf}
        \caption*{\tiny Figure 9: Theoretical (dashed) vs empirical (solid) $\Delta_k$ (p.12)}
    \end{figure}
\end{frame}

%\begin{frame}{Models--Datasets--Criteria}
%    \begin{center}
%        \begin{tabular}{|l|l|c|c|}
%            \hline
%            \textbf{Model} & \textbf{Dataset} & \boldmath$d$ & \boldmath$\Delta_{tol}$ \\
%            \hhline{|=|=|=|=|}
%            MLP (2 layers) & MNIST            & 10           & $1\times10^{-4}$        \\
%            \hline
%            MLP (2 layers) & Fashion-MNIST    & 10           & $1\times10^{-4}$        \\
%            \hline
%        \end{tabular}
%    \end{center}
%\end{frame}

\begin{frame}{Results and Conclusions}
    \begin{itemize}
        \item Identified threshold $k$ (MNIST), (Fashion-MNIST)
        \item Loss landscape stabilizes: additional data negligible beyond $k^*$
        \item Hessian-based bound provides reliable upper-bound for $\Delta_k$
        \item Practical guideline for dataset sizing and early stopping
    \end{itemize}
\end{frame}

>>>>

\documentclass{beamer}
\beamertemplatenavigationsymbolsempty
\usecolortheme{beaver}
\setbeamertemplate{blocks}[rounded=true, shadow=true]
\setbeamertemplate{footline}[page number]

\input{new_commands}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amssymb,amsfonts,amsmath,mathtext}
\usepackage{subfig}
\usepackage[all]{xy}
\usepackage{array}
\usepackage{multicol}
\usepackage{hyperref}
\usepackage{tabularx}   
\usepackage{hhline}

%----------------------------------------------------------------------------------
\title[Loss Landscape Convergence]{Neural Networks Loss Landscape Convergence in Hessian Low-Dimensional Space}
\author[Tem Nikitin \textit{et al.}]{Tem Nikitin}
\institute{Moscow Institute of Physics and Technology}
\date{2025}
%----------------------------------------------------------------------------------
\begin{document}

@@

\begin{frame}{Goals}
    \begin{itemize}
        \item Study how neural network loss landscape changes with dataset size
        \item Define and measure $\Delta_k = \mathbb{E}\bigl(\cL_{k+1}(\bw) - \cL_k(\bw)\bigr)^2$
        \item Develop Hessian-based low-dimensional projection method
        \item Derive theoretical bound for $\Delta_k$ via top-$d$ eigenvalues
        \item Propose an algorithm to determine the $\Delta$-sufficient dataset size
        \item Validate threshold $k^*$ beyond which further data yield negligible change
    \end{itemize}
\end{frame}

\begin{frame}{One-slide}
    \begin{columns}[c]
        \hspace{-1.5cm}
        \column{0.6\textwidth}
        \includegraphics[width=1.2\textwidth]{img/loss_eigen_-1_individ.pdf}
        \column{0.5\textwidth}
        \begin{itemize}
            \item Hessian projection onto top-$d$ eigenvectors
            \item Monte Carlo estimate of $\Delta_k^{emp} = \mathbb{E}\bigl(\cL_{k+1} - \cL_{k}\bigr)^2$
            \item Analytical bound: $\Delta_k^{th} \ge \Delta_k^{emp}$
            \item Detect dataset threshold $k^*$
        \end{itemize}
    \end{columns}
    \bigskip
    \textbf{Main message:} Loss landscape stabilizes after sufficient sample size.
\end{frame}

@@

\begin{frame}{Problem Statement: Quality Criteria}
    \begin{itemize}
        \item Convergence rate: $\Delta_k = O(1/k^2)$
        \item Theoretical bound via top-$d$ eigenvalues upper-bounds empirical $\Delta_k$
        \item Plateau in eigenvalue differences $\lambda_i^{k+1}-\lambda_i^k$ indicates threshold
    \end{itemize}
\end{frame}

\begin{frame}{Problem Solution: Theoretical Analysis}
    \begin{columns}[c]
        \column{0.4\textwidth}
        \begin{itemize}
            \item Project parameters: $\bw = \bw^* + \bP \btheta$, $\bP = [\be_1, \dots, \be_d]$
            \item Taylor approx: $\cL_k (\bw^* + \bP \btheta) \approx \cL_k (\bw^*) + \tfrac12 \btheta^{\T} \bLambda_k \btheta$
                  \bigskip
            \item Bound:
        \end{itemize}
        \column{0.6\textwidth}
        \includegraphics[width=1.0\textwidth]{img/eigenvalues.pdf}
        \vspace{0.5cm}\hspace{1.8cm}\raggedright{\tiny Eigenvalue decay}
    \end{columns}
    $$
        \Delta_k \approx
        \frac{\sigma^4}{4}\,\Biggl(2 \sum_{i=1}^d (\lambda_{k+1}^i - \lambda_k^i)^2
        + \Bigl(\sum_{i=1}^d (\lambda_{k+1}^i - \lambda_k^i)\Bigr)^2\Biggr).
    $$
\end{frame}

\begin{frame}{Goals of Computational Experiment}
    \begin{itemize}
        \item Datasets: MNIST, Fashion-MNIST (60k train, 10k test)
        \item MLP: 2 hidden layers, $\sim10^5$ parameters
        \item Subspace dimension $d=10$, Monte Carlo samples $K=64$, $\sigma=1$
        \item Compare empirical $\Delta_k$ vs theoretical bound across $k$
    \end{itemize}

    \begin{figure}
        \centering
        \includegraphics[width=1.05\textwidth,trim=0 17 0 1.5cm,clip]{img/delta_eigen_1_10_1024.pdf}
        \caption*{\tiny Figure 8: Monte Carlo $\Delta_k$ vs $k$ and $\Delta_k\cdot k^2$ (p.11)}
    \end{figure}
\end{frame}

@@

    \begin{figure}
        \centering
        \includegraphics[width=0.7\textwidth]{img/delta_border_1_10_1024.pdf}
        \caption*{\tiny Theoretical (dashed) vs empirical (solid) $\Delta_k$}
    \end{figure}
\end{frame}

\begin{frame}{Main Experiment: Theoretical vs. Empirical $\Delta_k$}
    \begin{figure}
        \centering
        \includegraphics[width=0.85\textwidth]{img/delta_loss_1_10_1024.pdf}
        \caption*{\tiny Empirical and Theoretical $\Delta_k$}
    \end{figure}
\end{frame}

\begin{frame}{Practical Measurements}
    \begin{table}
        \centering
        \small
        \begin{tabularx}{\textwidth}{|l|>{\raggedright\arraybackslash}X|c|c|c|c|}
            \hline
            \textbf{Dataset}
                          & \textbf{Model}    & \(\Delta\) & \(k\) & \(L_k\) & \textbf{Time (s)} \\
            \hhline{|=|=|=|=|=|=|}
            MNIST         & Single-layer MLP  & 0.025      & 100   & 0.013   & 2000              \\
                          & Multi-layer MLP   & 0.025      & 40    & 0.010   & 3500              \\
                          & Convolutional CNN & 0.025      & 60    & 0.024   & 1800              \\
            \hline
            Fashion-MNIST & Single-layer MLP  & 0.030      & 120   & 0.020   & 2100              \\
                          & Multi-layer MLP   & 0.030      & 90    & 0.017   & 4400              \\
                          & Convolutional CNN & 0.030      & 70    & 0.015   & 2400              \\
            \hline
        \end{tabularx}
    \end{table}
\end{frame}

\begin{frame}{Algorithm: \(\Delta\)-sufficient Dataset Size}
    \begin{block}{Algorithm 1: Determine \(\Delta\)-sufficient dataset size}
        \begin{enumerate}
            \item Initialize dataset \(D \gets \emptyset\) and batch counter \(k \gets 0\).
            \item \textbf{Repeat:}
                  \begin{itemize}
                      \item \(k \gets k + 1\).
                      \item Sample a new batch of data and append to \(D\).
                      \item Train or update model on \(D\).
                      \item Compute top-\(d\) Hessian eigenvalues \(\{\lambda_k^{(i)}\}_{i=1}^d\) and \(\{\lambda_{k+1}^{(i)}\}\).
                      \item Estimate
                            \[
                                \Delta_k \approx \frac{\sigma^4}{4}\Bigl(2\sum_{i=1}^d(\lambda_{k+1}^{(i)}-\lambda_k^{(i)})^2
                                + \bigl(\sum_{i=1}^d(\lambda_{k+1}^{(i)}-\lambda_k^{(i)})\bigr)^2\Bigr).
                            \]
                  \end{itemize}
            \item \textbf{Until} \(\Delta_k < \Delta_{\text{tol}}\).
            \item \textbf{Return} \(|D|\), the \(\Delta\)-sufficient sample size.
        \end{enumerate}
    \end{block}
\end{frame}

\begin{frame}{Results and Conclusions}
    \begin{itemize}
        \item Identified dataset thresholds \(k^*\) for MNIST and Fashion-MNIST.
        \item Loss landscape stabilizes: additional data negligible beyond \(k^*\).
        \item Hessian-based bound provides a reliable upper-bound for \(\Delta_k\).
        \item Proposed a practical algorithm for \(\Delta\)-sufficient dataset sizing.
        \item Offers guidelines for dataset collection and early stopping.
    \end{itemize}
\end{frame}

### 2025-05-22 slides/inn_prac/thesis/main.tex
Universal commits presents...

\documentclass[a4paper,12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{amsmath}
\geometry{margin=2.5cm}

@@

\noindent\textbf{UDC 519.688.4}\\[2ex]
\begin{center}
  {\LARGE \textbf{Neural Networks Loss Landscape Convergence in Hessian Low-Dimensional Space}}\\[2ex]
  Tem Nikitin$^{1}$, Nikita Kiselev$^{1}$, Andrey Grabovoy$^{1}$, Vladislav Meshkov$^{1}$\\[1ex]
  $^{1}$Moscow Institute of Physics and Technology, Moscow, Russia
\end{center}

@@

We tested on MNIST and Fashion-MNIST with simple MLP and CNN models. Figure~\ref{fig:delta} shows that the theoretical bound stays above the measured $\Delta_k$ curves and shrinks roughly like $O(1/k)$. Table~1 lists the minimal batch sizes $K$ needed for tolerances of 0.025 and 0.030. Values range from 40 to 120, and more complex models converge faster because they capture sharper directions.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.8\textwidth]{figure9.png}
  \caption{Empirical and theoretical $\Delta_k$ for MNIST experiments.}
  \label{fig:delta}
\end{figure}

\begin{table}[h]
  \centering
  \caption{Δ-sufficient sample size results}
  \begin{tabular}{lccc}
    \hline
    Dataset       & Model & $\Delta$ & $K$ \\
    \hline
    MNIST         & MLP   & 0.025    & 100 \\
    MNIST         & CNN   & 0.025    & 60  \\
    Fashion-MNIST & CNN   & 0.030    & 70  \\
    \hline
  \end{tabular}
\end{table}

>>>>

\documentclass[a4paper,12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\geometry{margin=2.5cm}

@@

\noindent\textbf{UDC 004.85}\\[2ex]
\begin{center}
  {\LARGE \textbf{Neural Networks Loss Landscape Convergence in Hessian Low-Dimensional Space}}\\[2ex]
  Tem Nikitin, Nikita Kiselev\\[1ex]
  Moscow Institute of Physics and Technology, Moscow, Russia
\end{center}
Understanding how a neural network’s loss landscape changes as we add more training data is important for efficient training. Although larger datasets reshape this high-dimensional surface, the point when extra data stop making a big difference is unclear. We show that near a local minimum the loss landscape stabilizes once the dataset exceeds a certain size. To study this, we project the full parameter space onto a lower-dimensional subspace formed by the Hessian’s top eigenvectors, highlighting the most important curvature directions. Within this subspace, we use Monte Carlo sampling to estimate how the loss changes more precisely. Experiments on standard image classification tasks demonstrate that our low-dimensional analysis pinpoints when the landscape stops evolving, offering practical guidance for balancing training cost with performance improvements.

@@

We tested on MNIST and Fashion-MNIST with simple MLP and CNN models. Figure~\ref{fig:delta} shows that the 
theoretical bound stays above the measured $\Delta_k$ curves and shrinks roughly like $O(1/k^2)$. Table~1 lists the minimal batch sizes $K$ needed for tolerances of 0.025 and 0.030. Values range from 40 to 120, and more complex models converge faster because they capture sharper directions.

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{img/delta_border_1_10_1024.pdf}
  \caption{Empirical and theoretical $\Delta_k$ for MNIST experiments.}
  \label{fig:delta}
\end{figure}

\begin{table}[ht]
  \centering
  \caption{$\Delta$-sufficient sample size results}
  \begin{tabular}{lccc}
    \hline
    Dataset & Model & $\Delta$ & $K$ \\
    \hline
    MNIST & Single-layer MLP & 0.025 & 100 \\
    MNIST & Convolutional CNN & 0.025 & 60 \\
    Fashion-MNIST & CNN & 0.030 & 70 \\
    \hline
  \end{tabular}
\end{table}
